---
title             : "Contextual cuing in the presence of an overt instruction"
shorttitle        : "Contextual cuing and instruction"
author: 
  - name          : "Tom Beesley"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, Lancaster University, UK, LA1 4YD"
    email         : "t.beesley@lancaster.ac.uk"
  - name          : "David Luque" 
    affiliation   : "2"
affiliation:
  - id            : "1"
    institution   : "Lancaster University, UK"
  - id            : "2"
    institution   : "Universidad Autónoma de Madrid, Spain"
abstract: |
  abstract here
  
  
  
  Public significance statement: 
keywords          : "keywords"
wordcount         : "X"
bibliography      : ["references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
header-includes:
  - \raggedbottom
  - \usepackage[font={small,it}, labelfont={bf}]{caption}
csl               : "apa7.csl"
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r knitr-options, echo=FALSE, message = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = FALSE, dev = "cairo_pdf")
```

```{r rmd-setup, include = FALSE}
library(tidyverse)
library(citr)
library(papaja)
library(english)
library(extrafont)
library(afex) 
library(BayesFactor)
library(apa)
library(pwr) # power calculations
library(lsr) # effect size estimate
# library(skimr)
# library(rstudioapi)
# library(kableExtra)

# function used to set rounding and presentation of decimals to specific length
spec_dec <- function(x, k) trimws(format(round(x, k), nsmall=k)) # useful for controlling decimals

# function to force scientific formatting of numbers (used for large BFs)
changeSciNot <- function(n) {
  output <- format(n, scientific = TRUE, digits = 2) #Transforms the number into scientific notation even if small
  output <- sub("e", "x10^", output) #Replace e with 10^
  output <- sub("\\+0?", "", output) #Remove + symbol and leading zeros on expoent, if > 1
  output <- sub("-0?", "-", output) #Leaves - symbol but removes leading zeros on expoent, if < 1
  output
}

# setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # set wd to current file location

theme_set(theme_classic())

#r_refs("references.bib")

```

Main text here [@beesley2015b]

```{r}

data_raw <- readRDS("combinedData.rds")

# get summary stats 

# how many participants in each experiment
data_raw %>%
  group_by(exp) %>% 
  summarise(num_Ps = n_distinct(subj))

# mean RT by participant
summary_rt <- 
  data_raw %>% 
  filter(timeout == 0,
         acc == 1) %>% 
  group_by(exp, subj) %>% 
  summarise(meanRT = mean(RT))

# plot as violin
summary_rt %>% 
  ggplot(aes(x = exp, y = meanRT)) +
  geom_violin(aes(fill = exp)) +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 2,
               binwidth = 25,
               alpha = .3)

# mean acc by participant
summary_acc <- 
  data_raw %>% 
  filter(timeout == 0) %>% 
  group_by(exp, subj) %>% 
  summarise(meanAcc = mean(acc))

# plot as violin
summary_acc %>% 
  ggplot(aes(x = exp, y = meanAcc)) +
  geom_violin(aes(fill = exp)) +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               binwidth = .01,
               dotsize = 1,
               alpha = .3)


# looking at emergence of CC in first few blocks across all experiments

data_raw %>% 
  filter(acc == 1,
         block %in% 1:6, 
         TT %in% c("random", "repeated")) %>%
  group_by(block, TT) %>% 
  summarise(meanRT = mean(RT, na.rm = TRUE)) %>% 
  ggplot(aes(x = block, colour = TT, y = meanRT, group = TT)) +
  geom_line(size = 1) +
  geom_point(size = 5) +
  theme_classic(base_size = 16)
  


```

# Experiment 1

Experiment 1 sought to examine whether the learnt attentional behaviour developed contextual cuing was expressed when participants were directed with a top-down instruction to search in a particular region of the search space. Participants were first trained with a set of four repeating configurations

## Method

### Participants

```{r}

# get basic stats for Ps

data_exp1 <- 
  data_raw %>% 
  filter(exp == "CCC01")

# rename TT variable
data_exp1 <- 
  data_exp1 %>% 
  mutate(TT = fct_recode(TT, 
                         "Repeated: consistent" = "repeated (C)",
                         "Repeated: inconsistent" = "repeated (I)",
                         "Random" = "random")) %>% 
  mutate(TT = fct_relevel(TT, 
                          "Repeated: consistent", "Repeated: inconsistent", "Random"))

demographics <- 
  data_exp1 %>% 
  group_by(subj) %>% 
  slice(1) %>% 
  select(subj, age, gender)

```

`r Words(nrow(demographics))` undergraduate students from Lancaster University were recruited (mean age = `r mean(demographics$age)`, SD = `r sd(demographics$age)`; `r count(demographics[,2:3], gender)[1,2]` identified as male and `r count(demographics[,2:3], gender)[2,2]` as female) via the Psychology Research Participation System in the Department of Psychology at Lancaster University, in return for the opportunity to use the recruitment system for their own research in future years. 

### Materials
Participants were tested individually in a quiet room with a Dell laptop with a 15.6” screen, a screen resolution of 1920 x 1080, and a full size external keyboard for participants to use to respond to the task. Participants sat approximately 50 cm from the screen. Stimulus presentation was controlled by MATLAB using the Psychophysics Toolbox extensions (Brainard, 1997; Kleiner, Brainard & Pelli, 2007; Pelli, 1997). Responses to the target stimulus were made by pressing the ‘c’ or ‘n’ key on a standard keyboard. All experimental materials are available at the github repository for this study. 

Distractor stimuli were an ‘L’ shape (rotated 0°, 90°, 180°, or 270°) while the target stimulus was a ‘T’ shape (rotated at either 90° or 270°). Stimuli were arranged in a square grid of 144 evenly spaced cells (12 x 12) which was positioned centrally on the screen and was XXX mm (XX°) square. The grid itself was invisible to participants. The fixation cross (displayed centrally before each trial) was XX mm (X.X°) square. The stimuli were XX mm (X.X°) square. The background of the screen was grey (RGB: .6, .6, .6) and the stimuli were presented in black. There was a small offset in the vertical line of the ‘L’ distractors, which increased the similarity between the ‘L’ distractor and the target ‘T’, making the search task more difficult (Duncan & Humphreys, 1989). 

### Design


### Procedure

## Results


```{r}

# participant exclusions 

to_summary <- 
  data_exp1 %>% 
  group_by(subj) %>% 
  summarise(meanTO = mean(timeout)) %>% 
  mutate(z_meanTO = scale(meanTO)[,1])

acc_summary <- 
  data_exp1 %>% 
  group_by(subj) %>% 
  summarise(meanAcc = mean(acc, na.rm = TRUE)) %>% 
  mutate(z_meanAcc = scale(meanAcc)[,1])

rt_summary <- 
  data_exp1 %>% 
  filter(acc == 1) %>% 
  group_by(subj) %>% 
  summarise(meanRT = mean(RT, na.rm = TRUE)) %>% 
  mutate(z_meanRT = scale(meanRT)[,1])

resp_summary_exp1 <- 
  to_summary %>% 
  left_join(acc_summary, by = "subj") %>% 
  left_join(rt_summary, by = "subj")

# is anyone an outlier in terms of timeouts?
outlier_to <- 
  resp_summary_exp1 %>% 
  filter(z_meanTO < -2.5 | z_meanTO > 2.5)

# is anyone an outlier in terms of accuracy?
outlier_acc <- 
  resp_summary_exp1 %>% 
  filter(z_meanAcc < -2.5 | z_meanAcc > 2.5)

# is anyone an outlier in terms of rt?
outlier_rt <- 
  resp_summary_exp1 %>% 
  filter(z_meanRT < -2.5 | z_meanRT > 2.5)

participants_removed <- 
  bind_rows(outlier_to, outlier_acc, outlier_rt) %>% 
  select(subj) %>% 
  mutate(subj = as.character(subj)) %>% 
  unique() %>% 
  pull()



```

Our criterion for removing outlier data, at both the participant level and the trial level, was 2.5 standard deviations above or below the mean of the sample. On average, trials ended with a timeout on `r spec_dec(100*mean(resp_summary_exp1$meanTO),2)`% of trials (SD = `r spec_dec(100*sd(resp_summary_exp1$meanTO),2)`). `r Words(nrow(outlier_to))` participants had an usually high proportion of timeouts and were removed from the analysis. The mean accuracy of participants (not including timeout trials) was `r spec_dec(100*mean(resp_summary_exp1$meanAcc),2)`% (SD = `r spec_dec(100*sd(resp_summary_exp1$meanAcc),2)`%). `r Words(nrow(outlier_acc))` participants that had an unusually low proportion of accurate trials and were also removed. The only participant deemed to be an outlier in terms of mean response time (hereafter RT) was also excluded on the basis of the timeout criterion, noted above. 

```{r}

data_exp1_OR <- 
  data_exp1 %>% 
  filter(!subj %in% participants_removed)

data_exp1_OR <- 
  data_exp1_OR %>% 
  filter(timeout == 0,
         acc == 1) %>% 
  group_by(subj) %>% 
  mutate(zRT = scale(RT)[,1]) # calculate z scores on participant basis

rt_outlier_summary <- 
  data_exp1_OR %>% 
  group_by(subj) %>% 
  summarise(prop_rem = mean(zRT < -2.5 | zRT > 2.5))

# was anyone an outlier in prop removed?
outlier_prop_removed <- 
  rt_outlier_summary %>% 
  mutate(z_prop_rem = scale(prop_rem)[,1]) %>% 
  filter(z_prop_rem < -2.5 | z_prop_rem > 2.5)

data_e1_f <-
  data_exp1_OR %>%
  ungroup() %>%
  filter(zRT > -2.5 & zRT < 2.5) # keep data in this range

```


For the remaining `r words(nrow(rt_outlier_summary))` participants we removed trials with a timeout and inaccurate trials, before removing outliers from the RT data. On average, the proportion of outliers removed was `r spec_dec(100*mean(rt_outlier_summary$prop_rem),2)`% (SD = `r spec_dec(100*sd(rt_outlier_summary$prop_rem),2)`%). `r words(nrow(outlier_prop_removed))` participants had an unusual proportion of trials removed as outlier RTs. 

Within-subject error bars were computed by a process of normalising the RT data for the sample [@cousineau2005]. Figure \@ref(fig:Exp1-RT-figure) shows the RT data across the 10 epochs of the experiment. In phase 1 (epochs 1-5) a contextual cuing effect rapidly emerged. In phase 2, the presence of the guiding arrow had a dramatic effect on the reduction of response times. Despite this, the underlying search configuration continued to play a role in the guidance of attention, with faster response times for (consistent) repeated configurations compared to random configurations.  

(ref:Exp1-RT-figure) RT data for Experiment 1.

```{r Exp1-RT-figure, fig.align="center", fig.cap="(ref:Exp1-RT-figure)", out.width="100%"}

# compute grand mean RT 
grand_mean_e1 <- mean(data_e1_f$RT)

# create subj level mean and normalised RT variable
data_e1_f <- 
  data_e1_f %>%
  group_by(subj) %>% 
  mutate(subj_mean_rt = mean(RT),
         normRT = RT - subj_mean_rt + grand_mean_e1) %>% 
  select(-subj_mean_rt)

# summarise the data across the 10 epochs, by subj
rt_summary_data_e1 <- 
  data_e1_f %>% 
  group_by(subj, epoch, TT) %>% 
  summarise(meanRT = mean(RT),
            mean_norm_RT = mean(normRT)) # used for normalised SE bars

# summarise at the group level and plot
rt_summary_data_e1 %>% 
  group_by(epoch, TT) %>% 
  summarise(group_RT = mean(meanRT),
            group_SE = sd(mean_norm_RT)/sqrt(n())) %>% 
  ggplot(aes(x = epoch, y = group_RT, group = TT, colour = TT)) +
  geom_line(size = .5) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = group_RT - group_SE,
                    ymax = group_RT + group_SE),
                width = .05)


```

```{r}
# Experiment 1 stats

# Statistics on Stage 1

phase_1 <- 
  rt_summary_data_e1 %>% 
  mutate(epoch = as.factor(epoch)) %>% 
  filter(epoch %in% c(1:5))

#exp1_phase1_anova <- aov_car(meanRT ~ Error(subj/TT*epoch), data = phase_1)

e1_phase1_BF <- 
  anovaBF(meanRT ~ TT * epoch + subj, 
          data = phase_1,
          whichRandom = "subj")

```

These data were explored with a Bayesian ANOVA, using the *BayesFactor::anovaBF()* function (for all analyses in this study the priors were set at the default "medium" width). First taking the data from phase 1 (epochs 1-5), the model with the largest Bayes Factor (BF) contained the factors of epoch and configuration (repeated vs. random), BF~10~ = `r changeSciNot(extractBF(e1_phase1_BF[3])$bf)`. The addition of the interaction term did not substantially improve the model predictions, BF~10~ `r spec_dec(extractBF(e1_phase1_BF[4])$bf/extractBF(e1_phase1_BF[3])$bf,1)`.

```{r}
# Experiment 1 stats

# Statistics on Stage 2

phase_2 <- 
  rt_summary_data_e1 %>% 
  mutate(epoch = as.factor(epoch)) %>% 
  filter(epoch %in% c(6:10))

#exp1_phase1_anova <- aov_car(meanRT ~ Error(subj/TT*epoch), data = phase_1)

e1_phase2_BF <- 
  anovaBF(meanRT ~ TT * epoch + subj, 
          data = phase_2,
          whichRandom = "subj")

# explore the main effect of configuration
phase_2_pairwise <- 
  data_e1_f %>% 
  mutate(epoch = as.factor(epoch)) %>% 
  filter(epoch %in% c(6:10)) %>%  
  group_by(subj, TT) %>% 
  summarise(meanRT = mean(RT))

# consistent vs. random
CvR <- 
  phase_2_pairwise %>% 
  filter(TT %in% c("Repeated: consistent", "Random")) 

anovaBF(meanRT ~ TT + subj, 
        data = CvR,
        whichRandom = "subj")

# inconsistent vs. random
IvR <- 
  phase_2_pairwise %>% 
  filter(TT %in% c("Repeated: inconsistent", "Random")) 

anovaBF(meanRT ~ TT + subj, 
        data = IvR,
        whichRandom = "subj")

# consistent vs. inconsistent 
CvI <- 
  phase_2_pairwise %>% 
  filter(TT %in% c("Repeated: consistent", "Repeated: inconsistent")) 

anovaBF(meanRT ~ TT + subj, 
        data = CvI,
        whichRandom = "subj")

```

A Bayesian ANOVA on the data from phase 2 (epochs 6-10) found significant support for the model containing the factor of configuration, BF~10~ = `r changeSciNot(extractBF(e1_phase2_BF[2])$bf)`. There was evidence to suggest that the addition of the factor of epoch did not substantially improve the model predictions, BF~10~ `r spec_dec(extractBF(e1_phase2_BF[3])$bf/extractBF(e1_phase2_BF[2])$bf,1)`.


```{r}
# Statistics on Stage 2

phase_2 <- 
  rt_summary_data_e1 %>% 
  mutate(epoch = as.numeric(epoch)) %>% 
  filter(epoch > 5)

aov_car(meanRT ~ Error(subj/TT*epoch), data = phase_2)

# random vs consistent
RvC <- 
  phase_2 %>% 
  filter(TT %in% c("random", "repeated (C)"))

t.test(meanRT ~ TT, data = RvC)

# random vs. inconsistent
RvI <- 
  phase_2 %>% 
  filter(TT %in% c("random", "repeated (I)"))

t.test(meanRT ~ TT, data = RvI)

# consistent vs. inconsistent
CvI <- 
  phase_2 %>% 
  filter(TT %in% c("repeated (C)", "repeated (I)"))

t.test(meanRT ~ TT, data = CvI)

```


# Experiment 2

Experiment 2 sought to examine ...

## Method

### Participants

```{r}

# get basic stats for Ps

data_exp2 <- 
  data_raw %>% 
  filter(exp == "CCC02") 

demographics <- 
  data_exp1 %>% 
  group_by(subj) %>% 
  slice(1) %>% 
  select(subj, age, gender)

```

`r Words(nrow(demographics))` undergraduate students from Lancaster University were recruited (mean age = `r mean(demographics$age)`, SD = `r sd(demographics$age)`; `r count(demographics[,2:3], gender)[1,2]` identified as male and `r count(demographics[,2:3], gender)[2,2]` as female) via the Psychology Research Participation System in the Department of Psychology at Lancaster University, in return for the opportunity to use the recruitment system for their own research in future years. 

### Materials
The materials and stimuli were identical to Experiment 1.

### Design


### Procedure

## Results


```{r}

# participant exclusions 

to_summary <- 
  data_exp2 %>% 
  group_by(subj) %>% 
  summarise(meanTO = mean(timeout)) %>% 
  mutate(z_meanTO = scale(meanTO)[,1])

acc_summary <- 
  data_exp2 %>% 
  group_by(subj) %>% 
  summarise(meanAcc = mean(acc, na.rm = TRUE)) %>% 
  mutate(z_meanAcc = scale(meanAcc)[,1])

rt_summary <- 
  data_exp2 %>% 
  filter(acc == 1) %>% 
  group_by(subj) %>% 
  summarise(meanRT = mean(RT, na.rm = TRUE)) %>% 
  mutate(z_meanRT = scale(meanRT)[,1])

resp_summary_exp2 <- 
  to_summary %>% 
  left_join(acc_summary, by = "subj") %>% 
  left_join(rt_summary, by = "subj")

# is anyone an outlier in terms of timeouts?
outlier_to <- 
  resp_summary_exp2 %>% 
  filter(z_meanTO < -2.5 | z_meanTO > 2.5)

# is anyone an outlier in terms of accuracy?
outlier_acc <- 
  resp_summary_exp2 %>% 
  filter(z_meanAcc < -2.5 | z_meanAcc > 2.5)

# is anyone an outlier in terms of rt?
outlier_rt <- 
  resp_summary_exp2 %>% 
  filter(z_meanRT < -2.5 | z_meanRT > 2.5)

participants_removed <- 
  bind_rows(outlier_to, outlier_acc, outlier_rt) %>% 
  select(subj) %>% 
  mutate(subj = as.character(subj)) %>% 
  unique() %>% 
  pull()

```

Our criteria for removing outlier data were identical to Experiment 1. On average, trials ended with a timeout on `r spec_dec(100*mean(resp_summary_exp2$meanTO),2)`% of trials (SD = `r spec_dec(100*sd(resp_summary_exp2$meanTO),2)`). `r Words(nrow(outlier_to))` participants had an usually high proportion of timeouts. The mean accuracy of participants (not including timeout trials) was `r spec_dec(100*mean(resp_summary_exp2$meanAcc),2)`% (SD = `r spec_dec(100*sd(resp_summary_exp2$meanAcc),2)`%). `r Words(nrow(outlier_acc))` participants that had an unusually low proportion of accurate trials and were also removed. `r Words(nrow(outlier_rt))` participants were deemed to be an outlier in terms of mean RT. 

```{r}

data_exp2_OR <- 
  data_exp2 %>% 
  filter(!subj %in% participants_removed)

data_exp2_OR <- 
  data_exp2_OR %>% 
  filter(timeout == 0,
         acc == 1) %>% 
  group_by(subj) %>% 
  mutate(zRT = scale(RT)[,1]) # calculate z scores on participant basis

rt_outlier_summary <- 
  data_exp2_OR %>% 
  group_by(subj) %>% 
  summarise(prop_rem = mean(zRT < -2.5 | zRT > 2.5))

# was anyone an outlier in prop removed?
outlier_prop_removed <- 
  rt_outlier_summary %>% 
  mutate(z_prop_rem = scale(prop_rem)[,1]) %>% 
  filter(z_prop_rem < -2.5 | z_prop_rem > 2.5)

data_e2_f <-
  data_exp2_OR %>%
  ungroup() %>%
  filter(zRT > -2.5 & zRT < 2.5) # keep data in this range

```


For the remaining `r words(nrow(rt_outlier_summary))` participants we removed trials with a timeout and inaccurate trials, before removing outliers from the RT data. On average, the proportion of outliers removed was `r spec_dec(100*mean(rt_outlier_summary$prop_rem),2)`% (SD = `r spec_dec(100*sd(rt_outlier_summary$prop_rem),2)`%). `r words(nrow(outlier_prop_removed))` participants had an unusual proportion of trials removed as outlier RTs and were not included in the final analysis. 

```{r}

# summarise the data across the 10 epochs, by subj
rt_summary_data_e2 <- 
  data_e2_f %>% 
  group_by(subj, patArrowP1, epoch, TT) %>% 
  summarise(meanRT = mean(RT))

# summarise at the group level and plot
rt_summary_data_e2 %>% 
  group_by(epoch, patArrowP1, TT) %>% 
  summarise(group_RT = mean(meanRT),
            group_SE = sd(meanRT)/sqrt(n())) %>% 
  ggplot(aes(x = epoch, y = group_RT, group = TT, colour = TT)) +
  geom_line(aes(group = interaction(TT, patArrowP1)), 
            size = .5) +
  geom_point(aes(shape = patArrowP1), 
             size = 5) +
  geom_errorbar(aes(ymin = group_RT - group_SE,
                    ymax = group_RT + group_SE),
                width = .05)


```

```{r}
# Statistics on Stage 1

phase_1 <- 
  rt_summary_data_e2 %>% 
  mutate(epoch = as.numeric(epoch)) %>% 
  filter(epoch <= 5)

aov_car(meanRT ~ Error(subj/patArrowP1*TT*epoch), data = phase_1)

# no evidence for an interaction of arrow training and TT
# is there evidence for the null?
phase_1_BF <- 
  anovaBF(meanRT ~ patArrowP1*TT + subj,
          data= data.frame(phase_1),
          whichRandom = "subj")

phase_1_BF[4]/phase_1_BF[3]
# the evidence favours an absence of an interaction effect

```

```{r}
# Statistics on Stage 2

phase_2 <- 
  rt_summary_data_e2 %>% 
  mutate(epoch = as.numeric(epoch)) %>% 
  filter(epoch > 5)

aov_car(meanRT ~ Error(subj/patArrowP1*TT*epoch), data = phase_2)

# no evidence for an interaction of arrow training and TT
# is there evidence for the null?
phase_2_BF <- 
  anovaBF(meanRT ~ patArrowP1*TT + subj,
          data= data.frame(phase_2),
          whichRandom = "subj")

phase_2_BF[4]/phase_2_BF[3]
# the evidence favours an absence of an interaction effect


```

# Experiment 3

Experiment 3 sought to examine ...

## Method

### Participants

```{r}

# get basic stats for Ps

data_exp3 <- 
  data_raw %>% 
  filter(exp == "CCC03") 

demographics <- 
  data_exp3 %>% 
  group_by(subj) %>% 
  slice(1) %>% 
  select(subj, age, gender)

```

`r Words(nrow(demographics))` undergraduate students from Lancaster University were recruited (mean age = `r mean(demographics$age)`, SD = `r sd(demographics$age)`; `r count(demographics[,2:3], gender)[1,2]` identified as male and `r count(demographics[,2:3], gender)[2,2]` as female) via the Psychology Research Participation System in the Department of Psychology at Lancaster University, in return for the opportunity to use the recruitment system for their own research in future years. 

### Materials
The materials and stimuli were identical to Experiment 1.

### Design


### Procedure

## Results


```{r}

# participant exclusions 

to_summary <- 
  data_exp3 %>% 
  group_by(subj) %>% 
  summarise(meanTO = mean(timeout)) %>% 
  mutate(z_meanTO = scale(meanTO)[,1])

acc_summary <- 
  data_exp3 %>% 
  group_by(subj) %>% 
  summarise(meanAcc = mean(acc, na.rm = TRUE)) %>% 
  mutate(z_meanAcc = scale(meanAcc)[,1])

rt_summary <- 
  data_exp3 %>% 
  filter(acc == 1) %>% 
  group_by(subj) %>% 
  summarise(meanRT = mean(RT, na.rm = TRUE)) %>% 
  mutate(z_meanRT = scale(meanRT)[,1])

resp_summary_exp3 <- 
  to_summary %>% 
  left_join(acc_summary, by = "subj") %>% 
  left_join(rt_summary, by = "subj")

# is anyone an outlier in terms of timeouts?
outlier_to <- 
  resp_summary_exp3 %>% 
  filter(z_meanTO < -2.5 | z_meanTO > 2.5)

# is anyone an outlier in terms of accuracy?
outlier_acc <- 
  resp_summary_exp3 %>% 
  filter(z_meanAcc < -2.5 | z_meanAcc > 2.5)

# is anyone an outlier in terms of rt?
outlier_rt <- 
  resp_summary_exp3 %>% 
  filter(z_meanRT < -2.5 | z_meanRT > 2.5)

participants_removed <- 
  bind_rows(outlier_to, outlier_acc, outlier_rt) %>% 
  select(subj) %>% 
  mutate(subj = as.character(subj)) %>% 
  unique() %>% 
  pull()

```

Our criteria for removing outlier data were identical to Experiment 1. On average, trials ended with a timeout on `r spec_dec(100*mean(resp_summary_exp3$meanTO),2)`% of trials (SD = `r spec_dec(100*sd(resp_summary_exp3$meanTO),2)`). `r Words(nrow(outlier_to))` participants had an usually high proportion of timeouts. The mean accuracy of participants (not including timeout trials) was `r spec_dec(100*mean(resp_summary_exp3$meanAcc),2)`% (SD = `r spec_dec(100*sd(resp_summary_exp3$meanAcc),2)`%). `r Words(nrow(outlier_acc))` participants that had an unusually low proportion of accurate trials and were also removed. `r Words(nrow(outlier_rt))` participants were deemed to be an outlier in terms of mean RT. 

```{r}

data_exp3_OR <- 
  data_exp3 %>% 
  filter(!subj %in% participants_removed)

data_exp3_OR <- 
  data_exp3_OR %>% 
  filter(timeout == 0,
         acc == 1) %>% 
  group_by(subj) %>% 
  mutate(zRT = scale(RT)[,1]) # calculate z scores on participant basis

rt_outlier_summary <- 
  data_exp3_OR %>% 
  group_by(subj) %>% 
  summarise(prop_rem = mean(zRT < -2.5 | zRT > 2.5))

# was anyone an outlier in prop removed?
outlier_prop_removed <- 
  rt_outlier_summary %>% 
  mutate(z_prop_rem = scale(prop_rem)[,1]) %>% 
  filter(z_prop_rem < -2.5 | z_prop_rem > 2.5)

data_e3_f <-
  data_exp3_OR %>%
  ungroup() %>%
  filter(zRT > -2.5 & zRT < 2.5) # keep data in this range

```


For the remaining `r words(nrow(rt_outlier_summary))` participants we removed trials with a timeout and inaccurate trials, before removing outliers from the RT data. On average, the proportion of outliers removed was `r spec_dec(100*mean(rt_outlier_summary$prop_rem),2)`% (SD = `r spec_dec(100*sd(rt_outlier_summary$prop_rem),2)`%). `r words(nrow(outlier_prop_removed))` participants had an unusual proportion of trials removed as outlier RTs and were not included in the final analysis [EAF4S]. 


```{r}

# summarise the data across the 10 epochs, by subj
rt_summary_data_e3 <- 
  data_e3_f %>% 
  group_by(subj, epoch, TT) %>% 
  summarise(meanRT = mean(RT))

# summarise at the group level and plot
rt_summary_data_e3 %>% 
  group_by(epoch, TT) %>% 
  summarise(group_RT = mean(meanRT),
            group_SE = sd(meanRT)/sqrt(n())) %>% 
  ggplot(aes(x = epoch, y = group_RT, group = TT, colour = TT)) +
  geom_line(size = .5) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = group_RT - group_SE,
                    ymax = group_RT + group_SE),
                width = .05)


```

```{r}
# Statistics on Stage 1

phase_1 <- 
  rt_summary_data_e3 %>% 
  mutate(epoch = as.numeric(epoch)) %>% 
  filter(epoch <= 5)

aov_car(meanRT ~ Error(subj/TT*epoch), data = phase_1)

```

```{r}
# Statistics on Stage 2

phase_2 <- 
  rt_summary_data_e3 %>% 
  mutate(epoch = as.numeric(epoch)) %>% 
  filter(epoch > 5)

aov_car(meanRT ~ Error(subj/TT*epoch), data = phase_2)

# Bayesian analysis of phase 2

anovaBF(meanRT ~ TT + subj,
        whichRandom = "subj",
        data = phase_2)

# repeated vs local random
d1v3 <- 
  phase_2 %>% 
  filter(TT %in% c("repeated","local (global random)"))

anovaBF(meanRT ~ TT + subj,
        whichRandom = "subj",
        data = d1v3)

# repeated vs global random
d1v4 <- 
  phase_2 %>% 
  filter(TT %in% c("repeated","global (local random)"))

anovaBF(meanRT ~ TT + subj,
        whichRandom = "subj",
        data = d1v4)

# local vs. random
d2v3 <- 
  phase_2 %>% 
  filter(TT %in% c("random","local (global random)"))

anovaBF(meanRT ~ TT + subj,
        whichRandom = "subj",
        data = d2v3)

# global vs. random
d2v4 <- 
  phase_2 %>% 
  filter(TT %in% c("random","global (local random)"))

anovaBF(meanRT ~ TT + subj,
        whichRandom = "subj",
        data = d2v4)

# local  vs. global  
d3v4 <- 
  phase_2 %>% 
  filter(TT %in% c("local (global random)","global (local random)"))

anovaBF(meanRT ~ TT + subj,
        whichRandom = "subj",
        data = d3v4)

t.test(meanRT ~ TT,
        data = d3v4,
       paired = TRUE)

```




\newpage

# References
