---
title             : "Contextual cuing in the presence of an overt instruction"
shorttitle        : "Contextual cuing and instruction"
author: 
  - name          : "Tom Beesley"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, Lancaster University, UK, LA1 4YD"
    email         : "t.beesley@lancaster.ac.uk"
  - name          : "David Luque" 
    affiliation   : "2"
affiliation:
  - id            : "1"
    institution   : "Lancaster University, UK"
  - id            : "2"
    institution   : "Universidad Autónoma de Madrid, Spain"

abstract: |
  abstract here
  
  
  Public significance statement: 
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
header-includes:
  - \raggedbottom
  - \usepackage[font={small,it}, labelfont={bf}]{caption}
csl               : "apa7.csl"
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r knitr-options, echo=FALSE, message = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = FALSE, dev = "cairo_pdf")
```

```{r rmd-setup, include = FALSE}
library(tidyverse)
library(citr)
library(papaja)
library(english)
# library(extrafont)
# library(afex) 
# library(BayesFactor)
# library(apa)
# library(pwr) # power calculations
# library(lsr) # effect size estimate
# library(skimr)
# library(rstudioapi)
# library(kableExtra)

# function used to set rounding and presentation of decimals to specific length
spec_dec <- function(x, k) trimws(format(round(x, k), nsmall=k)) # useful for controlling decimals

# function to force scientific formatting of numbers (used for large BFs)
changeSciNot <- function(n) {
  output <- format(n, scientific = TRUE, digits = 2) #Transforms the number into scientific notation even if small
  output <- sub("e", "x10^", output) #Replace e with 10^
  output <- sub("\\+0?", "", output) #Remove + symbol and leading zeros on expoent, if > 1
  output <- sub("-0?", "-", output) #Leaves - symbol but removes leading zeros on expoent, if < 1
  output
}

# setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # set wd to current file location

theme_set(theme_classic())

#r_refs("references.bib")

```

Main text here [@beesley2015b]

```{r}

data_raw <- readRDS("combinedData.rds")

# get summary stats 

# how many participants in each experiment
data_raw %>%
  group_by(exp) %>% 
  summarise(num_Ps = n_distinct(subj))

# mean RT by participant
summary_rt <- 
  data_raw %>% 
  filter(timeout == 0,
         acc == 1) %>% 
  group_by(exp, subj) %>% 
  summarise(meanRT = mean(RT))

# plot as violin
summary_rt %>% 
  ggplot(aes(x = exp, y = meanRT)) +
  geom_violin(aes(fill = exp)) +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 2,
               binwidth = 25,
               alpha = .3)

# mean acc by participant
summary_acc <- 
  data_raw %>% 
  filter(timeout == 0) %>% 
  group_by(exp, subj) %>% 
  summarise(meanAcc = mean(acc))

# plot as violin
summary_acc %>% 
  ggplot(aes(x = exp, y = meanAcc)) +
  geom_violin(aes(fill = exp)) +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               binwidth = .01,
               dotsize = 1,
               alpha = .3)


```

# Experiment 1

Experiment 1 sought to examine whether the learnt attentional behaviour developed contextual cuing was expressed when participants were directed with a top-down instruction to search in a particular region of the search space. Participants were first trained with a set of four repeating configurations

## Method

### Participants

```{r}

# get basic stats for Ps

data_exp1 <- 
  data_raw %>% 
  filter(exp == "CCC01") 

demographics <- 
  data_exp1 %>% 
  group_by(subj) %>% 
  slice(1) %>% 
  select(subj, age, gender)

```

`r Words(nrow(demographics))` undergraduate students from Lancaster University were recruited (mean age = `r mean(demographics$age)`, SD = `r sd(demographics$age)`; `r count(demographics[,2:3], gender)[1,2]` identified as male and `r count(demographics[,2:3], gender)[2,2]` as female) via the Psychology Research Participation System in the Department of Psychology at Lancaster University, in return for the opportunity to use the recruitment system for their own research in future years. 

### Materials
Participants were tested individually in a quiet room with a Dell laptop with a 15.6” screen, a screen resolution of 1920 x 1080, and a full size external keyboard for participants to use to respond to the task. Participants sat approximately 50 cm from the screen. Stimulus presentation was controlled by MATLAB using the Psychophysics Toolbox extensions (Brainard, 1997; Kleiner, Brainard & Pelli, 2007; Pelli, 1997). Responses to the target stimulus were made by pressing the ‘c’ or ‘n’ key on a standard keyboard. All experimental materials are available at the github repository for this study. 

Distractor stimuli were an ‘L’ shape (rotated 0°, 90°, 180°, or 270°) while the target stimulus was a ‘T’ shape (rotated at either 90° or 270°). Stimuli were arranged in a square grid of 144 evenly spaced cells (12 x 12) which was positioned centrally on the screen and was XXX mm (XX°) square. The grid itself was invisible to participants. The fixation cross (displayed centrally before each trial) was XX mm (X.X°) square. The stimuli were XX mm (X.X°) square. The background of the screen was grey (RGB: .6, .6, .6) and the stimuli were presented in black. There was a small offset in the vertical line of the ‘L’ distractors, which increased the similarity between the ‘L’ distractor and the target ‘T’, making the search task more difficult (Duncan & Humphreys, 1989). 

### Design


### Procedure

## Results


```{r}

# participant exclusions 

to_summary <- 
  data_exp1 %>% 
  group_by(subj) %>% 
  summarise(meanTO = mean(timeout)) %>% 
  mutate(z_meanTO = scale(meanTO)[,1])

acc_summary <- 
  data_exp1 %>% 
  group_by(subj) %>% 
  summarise(meanAcc = mean(acc, na.rm = TRUE)) %>% 
  mutate(z_meanAcc = scale(meanAcc)[,1])

rt_summary <- 
  data_exp1 %>% 
  filter(acc == 1) %>% 
  group_by(subj) %>% 
  summarise(meanRT = mean(RT, na.rm = TRUE)) %>% 
  mutate(z_meanRT = scale(meanRT)[,1])

resp_summary_exp1 <- 
  to_summary %>% 
  left_join(acc_summary, by = "subj") %>% 
  left_join(rt_summary, by = "subj")

# is anyone an outlier in terms of timeouts?
outlier_to <- 
  resp_summary_exp1 %>% 
  filter(z_meanTO < -2.5 | z_meanTO > 2.5)

# is anyone an outlier in terms of accuracy?
outlier_acc <- 
  resp_summary_exp1 %>% 
  filter(z_meanAcc < -2.5 | z_meanAcc > 2.5)

# is anyone an outlier in terms of rt?
outlier_rt <- 
  resp_summary_exp1 %>% 
  filter(z_meanRT < -2.5 | z_meanRT > 2.5)

participants_removed <- 
  bind_rows(outlier_to, outlier_acc, outlier_rt) %>% 
  select(subj) %>% 
  mutate(subj = as.character(subj)) %>% 
  unique() %>% 
  pull()

```

Our criterion for removing outlier data, at both the participant level and the trial level, was 2.5 standard deviations above or below the mean of the sample. On average, trials ended with a timeout on `r spec_dec(100*mean(resp_summary_exp1$meanTO),2)`% of trials (SD = `r spec_dec(100*sd(resp_summary_exp1$meanTO),2)`). `r Words(nrow(outlier_to))` participants had an usually high proportion of timeouts and were removed from the analysis. The mean accuracy of participants (not including timeout trials) was `r spec_dec(100*mean(resp_summary_exp1$meanAcc),2)`% (SD = `r spec_dec(100*sd(resp_summary_exp1$meanAcc),2)`%). `r Words(nrow(outlier_to))` participants had an usually low proportion of accurate trials and were also removed. The only participant deemed to be an outlier in terms of mean response time (hereafter RT) was also excluded on the basis of the timeout criterion, noted above. 

```{r}

data_exp1_OR <- 
  data_exp1 %>% 
  filter(!subj %in% participants_removed)

data_exp1_OR <- 
  data_exp1_OR %>% 
  filter(timeout == 0,
         acc == 1) %>% 
  group_by(subj) %>% 
  mutate(zRT = scale(RT)[,1]) # calculate z scores on participant basis

rt_outlier_summary <- 
  data_exp1_OR %>% 
  group_by(subj) %>% 
  summarise(prop_rem = mean(zRT < -2.5 | zRT > 2.5))

# was anyone an outlier in prop removed?
rt_outlier_summary %>% 
  mutate(z_prop_rem = scale(prop_rem)[,1]) %>% 
  filter(z_prop_rem < -2.5 | z_prop_rem > 2.5)

data_e1_f <-
  data_exp1_OR %>%
  ungroup() %>%
  filter(zRT > -2.5 & zRT < 2.5) # keep data in this range

```


For the remaining `r words(nrow(rt_outlier_summary))`participants we removed trials with a timeout and inaccurate trials, before removing outliers from the RT data. On average, the proportion of outliers removed was `r spec_dec(100*mean(rt_outlier_summary$prop_rem),2)`% (SD = `r spec_dec(100*sd(rt_outlier_summary$prop_rem),2)`%). No participants had an unusual proportion of trials removed as outlier RTs. 

```{r}

# summarise the data across the 10 epochs, by subj
rt_fig_summary <- 
  data_e1_f %>% 
  group_by(subj, epoch, TT) %>% 
  summarise(meanRT = mean(RT))

# compute normalised RT (norm_RT), for calculating w/s error bars (i.e., Cosin)
rt_fig_summary <- 
  rt_fig_summary %>% 
  group_by(subj) %>% # within each experiment, and each subject
  summarise(sub_MeanRT = mean(meanRT)) %>% # calculate the mean RT for the subject
  mutate(sample_MeanRT = mean(sub_MeanRT)) %>% # the overall mean RT for the sample in that experiment
  ungroup() %>% # ungroup the data, to allow for removal of variable
  left_join(rt_fig_summary, by = "subj") %>% # join to the original data frame
  mutate(norm_RT = meanRT - sub_MeanRT + sample_MeanRT) # compute the normalised RT, from the RT, subject mean and sample mean



# summarise at the group level and plot
rt_fig_summary %>% 
  group_by(epoch, TT) %>% 
  summarise(group_RT = mean(meanRT),
            group_SE = sd(norm_RT)/sqrt(n())) %>% 
  ggplot(aes(x = epoch, y = group_RT, group = TT, colour = TT)) +
  geom_line(size = .5) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = group_RT - group_SE,
                    ymax = group_RT + group_SE),
                width = .05)


```


\newpage

# References
